---
title: "BREP Report"
author: "Tomo Eguchi"
date: "August 30, 2018"
output: word_document
---

```{r setup, include=FALSE, echo=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(dplyr)
library(ggplot2)

save.data <- T
save.fig <- T

ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))

get.lat <- function(x){
  lat.deg <- as.numeric(paste(x[21:22], sep = '', collapse = ''))
  lat.min <- as.numeric(paste(x[24:25], sep = '', collapse = ''))
  lat.min <- lat.min + as.numeric(paste(x[27:28],
                                        sep = '', collapse = ''))/100
  lat <- lat.deg + lat.min/60
  return(lat)

}

get.lon <- function(x){
  lon.deg <- as.numeric(paste(x[31:33], sep = '', collapse = ''))
  lon.min <- as.numeric(paste(x[35:36], sep = '', collapse = ''))
  lon.min <- lon.min + as.numeric(paste(x[38:39],
                                        sep = '', collapse = ''))/100
  lon <- -1 * (lon.deg + lon.min/60)
  return(lon)
}
```

```{r get_data, include=F, echo=F, cache=TRUE}
infiles <- c('Data/das files/SCB2018_DAS_ed1.822',
             "Data/das files/DAS1235_ed5.825",
             "Data/das files/SCB2018_DAS_ed1.826",
             "Data/das files/DAS1116_ed2.828",
             "Data/das files/DAS1001_ed1.829")

outfile <- 'Data/tmpTracks_2018.txt'
Lon1 <- Lat1 <- Time1 <- Date1 <-  NA
Line <- Distance <- Lon2<-Lat2 <- Time2<-Date2 <-  NA
Lon.cc <- Lat.cc <- Date.cc <- Time.cc <- NA
c <- cc <- 1

for (k1 in 1:length(infiles)){
  y <- readLines(infiles[k1])
  k <- 1
  
  for(k in 1:length(y)){
    x <- unlist(strsplit(y[k], ''))
    
    lineNum <- as.numeric(paste(x[1:3], sep = '', collapse = ''))
    eventCode <- x[4]
    if (eventCode == 'T' | eventCode == 'R'){
      Date1[c] <- paste(x[13:18], sep = '', collapse = '')
      Time1[c] <- paste(x[6:11], sep = '', collapse = '')
      Lat1[c] <- get.lat(x)
      Lon1[c] <- get.lon(x)
      line.tmp <- paste(na.omit(x[40:44]), sep = '', collapse = '')
      line.tmp <- as.numeric(line.tmp)
      Line[c] <- ifelse(!is.na(line.tmp), as.numeric(line.tmp), Line[c-1])
      
    } else if (eventCode == 'O' | eventCode == 'E') {
      Date2[c] <- paste(x[13:18], sep = '', collapse = '')
      Time2[c] <- paste(x[6:11], sep = '', collapse = '')
      Lat2[c] <- get.lat(x)
      Lon2[c] <- get.lon(x)
      c <- c + 1
    } else if (eventCode == "t"){
      Date.cc[cc] <- paste(x[13:18], sep = '', collapse = '')
      Time.cc[cc] <- paste(x[6:11], sep = '', collapse = '')
      Lat.cc[cc] <- get.lat(x)
      Lon.cc[cc] <- get.lon(x)
      cc <- cc + 1
    }
    
  }
  
}

Distance <- km2nm(geosphere::distGeo(cbind(Lon1, Lat1),
                                     cbind(Lon2, Lat2))/1000)

out.df <- na.omit(data.frame(Date1 = Date1, Time1 = Time1, Lat1 = Lat1, Lon1 = Lon1,
                     Date2 = Date2, Time2 = Time2, Lat2 = Lat2, Lon2 = Lon2,
                     Line = Line, Distance = Distance))

cc.df <- data.frame(Date = Date.cc, Time = Time.cc, Lat = Lat.cc, Lon = Lon.cc)

if(save.fig)
  write.csv(out.df, file = outfile,
            quote = F, row.names = F)
```

```{r get_lines, echo=F, include=F, cache=TRUE}

# These track lines were created for the BREP funded study 
track.lines <- read.csv("Data/tracklines_SC_N_10km_2017-08-27.csv",
                        header = T)

study.area <- readRDS(file = "RDSfiles/SearchArea_2018-08-24.rds")
coast.line <- readRDS(file = "RDSfiles/CoastLine_2018-08-24.rds")
transect.lines <- readRDS(file = "RDSfiles/TransectLines_2018-08-24.rds") %>%
  filter(lineID != 12 & lineID != 14)

offshore.extensions <- read.csv(file = "Data/offshoreSegments.csv",
                                header = T)
MX.lines <- read.csv(file = "Data/MX_Segments.csv",
                           header = T)
```

```{r plot, echo=F, include=F}
land.color <- '#333333'

p.1 <- ggplot() +
  geom_polygon(fill = land.color,
               data = coast.line,
               aes(x = X, y = Y, group = idx))  +
  # geom_polygon(data = study.area,
  #              aes(x = X, y = Y),
  #              fill = NA, size = 1.2,
  #              alpha = 0.6,
  #              color='black') +
  # geom_segment(data = transect.lines,
  #              aes(x = lonInshore, y = latInshore,
  #                  xend = lonOffshore, yend = latOffshore),
  #              size = 1.0) +
  # geom_text(data = transect.lines,
  #           aes(x = lonOffshore,
  #               y = latOffshore, label = lineID),
  #           color = 'firebrick1',
  #           fontface = "bold")+
  # geom_segment(data = offshore.extensions,
  #              aes(x = lon1, y = lat1,
  #                  xend = lon2, yend = lat2),
  #              size = 1.0) +
  # geom_segment(data = MX.lines,
  #              aes(x = lon1, y = lat1,
  #                  xend = lon2, yend = lat2),
  #              size = 1.0) +
  # geom_text(data = offshore.extensions,
  #           aes(x = lon2,
  #               y = lat2, label = lineID),
  #           color = 'firebrick1',
  #           fontface = "bold")+
  geom_segment(data = out.df,
               aes(x = Lon1, y = Lat1,
                   xend = Lon2, yend = Lat2),
               color = "firebrick",
               alpha = 0.4,
               size = 1.2) +
  geom_point(data = cc.df,
             aes(x = Lon, y = Lat),
             color = "red",
             size = 2) +
  coord_map() +
  xlab('') + ylab('')

if (save.fig)
  ggsave(filename = paste0("figures/finished_tracklines_BREP_",
                           Sys.Date(), ".png"),
         plot = p.1, dpi = 600, device = "png")

```

Aerial survey was conducted between 22 and 29 August 2018 in the Southern California Bight (Figure 2). The survye was conducted by Tomo Eguchi and colleagues from NOAA Southwest Fisheries Science Center aboard a hihg-wing twin-engine aircraft (Partenavia Observer). The aircraft and a pilot were contracted from Aspen Helicopters (Oxnard, CA). Coastal clouds, high winds, and military exercises were main obstacles for the survey. The team completed approximately  ```r round(sum(out.df$Distance))``` nm of track lines on five survey days. During the survey, two loggerhead turtles were sighted (Figure 2). 

```{r plot_tracks, echo=FALSE, cache=TRUE, fig.cap = "Figure 2. Completed track lines of the SCB aerial survey for loggerhead turtles. Two red dots indicate loggerhead turtle sightings."}
knitr::include_graphics(paste0("figures/finished_tracklines_BREP_", Sys.Date(), ".png"))
```
