---
title: "SCB E-W transect lines"
output: html_notebook
---

This document describes drawing transect lines within the SCB for marine mammal surveys.

Adapted from the same named file in the SDB project.

First load necessary libraries.
```{r}
rm(list=ls())
save.image <- F
save.data <- F

# should move to sf package.
library(rgdal)
library(sp)
library(geosphere)
library(rgeos)
library(tidyverse)
library(sf)

## Convert nm to km
nm2km <- function(nm){
  km <- nm * 1.852
  return(km)
}

# convert km to nm
km2nm <- function(km){
  nm <- km / 1.852
  return(nm)
}

# length of each line from the coast line (m)
line.length <- nm2km(20) * 1000 

dlines <- nm2km(2) * 1000    # distance between lines in m
min.length <- nm2km(1) * 1000  # minimum line length in m

water.color <- "lightblue"
background.color <- "darkgray"
eelgrass.color <- "lightgreen"


#getCoastLine
# A function to extract a coast line given longitude and latitude
# limits for the Pacific Ocean

# Data files - E and W Pacific are separated; got these from Rich Cosgrove

# Used to be in Matlab

# Tomo Eguchi
# 16 December 2016
getCoastLine <- function(filename, lon.limits, lat.limits){
  data1 <- read.table(filename, header = F, row.names = NULL, sep = ',')
  # fix longitudes to -180 to 180
  lon.limits[lon.limits > 180] <- lon.limits[lon.limits > 180] - 360
  out.df <- data1[data1[,1] >= min(lon.limits) &
                    data1[,1] <= max(lon.limits) &
                    data1[,2] >= min(lat.limits) &
                    data1[,2] <= max(lat.limits),]
  colnames(out.df) <- c('Longitude', 'Latitude')
  out.df$idx <- NA
  idx.rows <- 1:nrow(out.df)
  na.idx <- is.na(out.df$Longitude)  # T/F for NA
  dif.na.idx <- na.idx[2:length(na.idx)] - na.idx[1:(length(na.idx)-1)]
  idx.neg1 <- idx.rows[dif.na.idx == -1]  # beginning of segments
  idx.pos1 <- idx.rows[dif.na.idx == 1]   # end of segments
  
  for (k in 1:length(idx.neg1)) {
    out.df[idx.neg1[k]:idx.pos1[k], 'idx'] <- k
  }
  
  # change index to a factor variable
  out.df$idx <- as.factor(out.df$idx)
  out.df <- na.omit(out.df)
  # this splits all segments into separate dataframes
  out.list <- split(out.df, out.df$idx)
  
  for (k in 1:length(out.list)){
    line.1 <- out.list[[k]][1,]
    line.end <- out.list[[k]][nrow(out.list[[k]]),]
    if (line.1$Longitude != line.end$Longitude){
      tmp <- out.list[[k]]
      tmp <- rbind(c(max(tmp$Longitude), max(tmp$Latitude), k),
                   tmp, c(max(tmp$Longitude), max(tmp$Latitude), k))
      out.list[[k]] <- tmp
    }
  }
  return(out.list)
  
}

home.dir <- Sys.getenv("HOME")

if (length(grep("Documents", home.dir)) != 0){
  home.dir <- strsplit(home.dir, "Documents")[[1]]
} else {
  home.dir <- home.dir
}

fishery.closure.2021.10.08 <- data.frame(X = c(-(118+6/60), 
                                               -(118+10/60), 
                                               -(117+54.5/60), 
                                               -(117+44.75/60),
                                               -(117+32.25/60)),
                                         Y = c(33+44.25/60,
                                               33+38/60,
                                               33+11/60,
                                               33+11/60,
                                               33+21.75/60))

fishery.closure.2021.10.08.sf <- st_as_sf(fishery.closure.2021.10.08, 
                                          coords = c("X", "Y"), 
           crs = "+proj=longlat +datum=WGS84") 

coast.line.all <- getCoastLine(paste0(home.dir, 
                                  "Documents/Oceans and Maps/coast/coast_Epac.txt"),
                           c(-120.7, -116.6),
                           c(30.98, 34.58))

coast.line <- coast.line.all[[1]]

approx.center <- data.frame(X=-119.967, Y=33.092)
cetner.UTM <- st_as_sf(approx.center, coords = c("X", "Y"), 
           crs = "+proj=longlat +datum=WGS84") #%>%
   #st_transform(crs = "+proj=utm +zone=10 ellps=WGS84")
  
# N/S most latitudes for transect lines
N.most.lat <- 33.73    # Seal Beach 
#N.most.lat <- 33.628  # approx Huntington Beach Pier
#S.most.lat <- 33.210  # approx Oceanside Harbor
S.most.lat <- 32.67    # Cabrillo National Monument

#NW.corner <- c(-118.3362, 33.4639)  # create a NW corner (for Huntington Beach Pier)
NW.corner <- c(-118.38, 33.5) # for Seal Beach

E.end <- -117
W.end <- NW.corner[1]  # -118.3  # approx S tip of Catalina


#load('RData/studyArea.RData')

dpi <- 600
```

The following code was modified from a script, which was used to draw aerial transect lines for the Southern California Bight (CaptureArea.R in SCB_AerialSurvey project).


First line is defined by selecting the northern most latitude for transect lines and finding the longitude of the shore lines (both E and W). This is completed by using gIntersection.  The while loop continues until the new latitude is south of the southern limit. 
```{r}

# get county lines
county.lines <- st_read(paste0(home.dir, 
                               "Documents/Oceans and Maps/CA_Counties"))

county.lines %>% filter(NAME == "San Diego") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") -> SD.county.line

county.lines %>% filter(NAME == "Los Angeles") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") -> LA.county.line

county.lines %>% filter(NAME == "Orange") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") -> OR.county.line

search.area <- filter(coast.line, 
                      Latitude <= N.most.lat & Latitude >= S.most.lat) %>%
  select("Longitude", "Latitude")

coast.line.linestring <- st_as_sf(coast.line, 
                                  coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") %>%
  st_combine() %>%
  st_cast("LINESTRING")

N.most <- filter(search.area, Latitude == max(search.area$Latitude))
S.most <- filter(search.area, Latitude == min(search.area$Latitude))

# using the northern boundary as the guide, we find the rhumb line:
rhumb_bear <- bearingRhumb(N.most, NW.corner)

# create a line along this latitude (longitude values are arbitrary):
N.most.line <- st_as_sf(data.frame(X = c(W.end, E.end),
                                   Y = c(N.most.lat, N.most.lat)),
                        coords = c("X", "Y"),
                        crs = "+proj=longlat +datum=WGS84") %>%
  st_combine() %>%
  st_cast("LINESTRING")

# E end of the northern most line - intersection with the coast line
N.coords <- st_intersection(coast.line.linestring, N.most.line) %>%
  st_cast("POINT")

N.coords.df <- st_coordinates(N.coords) %>% 
  data.frame() %>% 
  filter(X == max(X))

# find the end point 
N.coords.df[2,] <- destPoint(N.coords.df, rhumb_bear, line.length)

# set up output dataframes
coords.all <- data.frame(X = NA, Y = NA, ID = NA)
E.coords <- data.frame(long = N.coords.df[1, "X"],
                       lat = N.coords.df[1, "Y"])

W.coords <- data.frame(long = N.coords.df[2, "X"],
                       lat = N.coords.df[2, "Y"])

lat <- E.coords$lat

k <- k1 <- 1
while (lat > S.most.lat){
  # find new latitude - 90 degrees off from the previous line
  tmp.pt <- data.frame(destPoint(E.coords[k, ], rhumb_bear-90, dlines))
  lat <- tmp.pt[2]
  # tmp.pt is not on the coast line find the arbitrary point offshore 
  # along the new line:
  tmp.coords <- destPoint(tmp.pt, rhumb_bear, line.length)
  
  # create a point on the land 3nm is arbitrary
  tmp.coords.onland <- destPoint(tmp.coords, 
                                 rhumb_bear-180, 
                                 line.length + (nm2km(3)*1000))
  
  # create a temporary line from the land to tmp.coords
  tmp.line <-  st_as_sf(data.frame(X = c(tmp.coords[1,1], 
                                         tmp.coords.onland[1,1]),
                                   Y = c(tmp.coords[1,2],
                                         tmp.coords.onland[1,2])),
                        coords = c("X", "Y"),
                        crs = "+proj=longlat +datum=WGS84") %>%
    st_combine() %>%
    st_cast("LINESTRING")
  
  # find the intersection between tmp.line and coast line
  inshore.pt <- st_intersection(coast.line.linestring, tmp.line) %>%
    st_cast("POINT") %>% 
    st_coordinates() %>%
    data.frame() %>%
    filter(X == max(X))
  
  # find the point 20 nm on rhumb_bear direction
  offshore.pt <- destPoint(inshore.pt, rhumb_bear, line.length)

  new.line <-  st_as_sf(data.frame(X = c(inshore.pt[1,1],
                                         offshore.pt[1,1]),
                                   Y = c(inshore.pt[1,2],
                                         offshore.pt[1,2])),
                        coords = c("X", "Y"),
                        crs = "+proj=longlat +datum=WGS84") %>%
    st_combine() %>%
    st_cast("LINESTRING")
  
  new.coords.df <- st_coordinates(new.line)
  
  coords.all[k1,] <- cbind(new.coords.df[1,], k)
  coords.all[k1+1,] <- cbind(new.coords.df[2,], k)
  k1 <- k1 + 2
  k <- k + 1
  W.coords[k, ] <- new.coords.df[2, 1:2]
  E.coords[k, ] <- new.coords.df[1, 1:2]

}

# I can't figure out how to add crs once a multilinesting object is
# created... for a later date. Wasted enough time 2021-10-08
# multipoints <- st_as_sf(coords.all,
#                         coords = c("X", "Y"),
#                         crs = "+proj=longlat +datum=WGS84") %>%
#   st_cast("MULTIPOINT",
#           crs = "+proj=longlat +datum=WGS84")
# 
# points <- st_cast(st_geometry(multipoints), "POINT",
#                         crs = "+proj=longlat +datum=WGS84")
# n <- max(coords.all$ID)
# linestrings <- lapply(X = 1:n, FUN = function(x) {
#   
#   pair <- st_combine(c(points[x], points[x + 1]))
#   line <- st_cast(pair, "LINESTRING",
#                   crs = "+proj=longlat +datum=WGS84")
#   return(line)
# 
# })
# 
# multilinestrings <- st_multilinestring(do.call("rbind", linestrings)) 
# 
# tmp <- st_transform(st_multilinestring(multilinestrings), 
#                     "+proj=longlat +datum=WGS84")

W.coords %>% transmute(long.W = long, lat.W = lat) -> W.coords
E.coords %>% transmute( long.E = long, lat.E = lat) -> E.coords

# find distance between beginning and end:
dist.line <- distHaversine(W.coords, E.coords)

# get rid of those that are < min.length m, then reorder lat/lon so that I can copy and paste
# in Google Earth to plot end points easily (they like lat, long pairs, not long, lat)

all.coords <- cbind(W.coords, E.coords, dist.line)
all.coords %>% filter(dist.line > min.length) %>%
  filter(lat.W > S.most.lat) %>%
  transmute(lat.W = lat.W, long.W = long.W,
            lat.E = lat.E, long.E = long.E,
            length = dist.line) -> survey.coords

survey.coords$ID <- seq(1, nrow(survey.coords))

survey.coords %>% select(ID, lat.W, long.W, lat.E, long.E) -> survey.coords
```

Create another output data.frame

```{r}
survey.coords.2 <- data.frame(Pt = as.character(),
                              Lat = as.double(),
                              Lon = as.double())
c <- 1
for (k in 1:nrow(survey.coords)){
  survey.coords.2[c,1] <- paste0(survey.coords[k,1], "E")
  survey.coords.2[c,2:3] <- c(survey.coords[k, "lat.E"], 
                              survey.coords[k, "long.E"])
  survey.coords.2[c+1,1] <- paste0(survey.coords[k,1], "W")
  survey.coords.2[c+1,2:3] <- c(survey.coords[k, "lat.W"], 
                             survey.coords[k, "long.W"])
  c <- c + 2  
}


```


Once W and E ends are created, plot them on a map.

```{r}
# water.gis <- spTransform(readOGR(dsn = "GISfiles",
#                                  layer = "water",
#                                  verbose = FALSE),
#                          CRS("+proj=longlat +datum=WGS84"))
# 
# water.df <- broom::tidy(water.gis)

coast.line.all.df <- do.call("rbind", coast.line.all) %>%
  group_by(idx)
# make a plot
p.1 <- ggplot() +
  geom_polygon(data = coast.line.all.df,
            aes(x = Longitude, y = Latitude, group = idx),
            fill = background.color, color = "black") +
  geom_path(data = fishery.closure.2021.10.08,
            aes(x = X, y = Y), color = "darkblue") +
  geom_segment(data = survey.coords,
               aes(x = long.W, y = lat.W,
                   xend = long.E, yend = lat.E),
               size = 0.1) +
  theme(panel.background = element_rect(fill = water.color)) +
  geom_text(data = survey.coords,
            aes(x = long.W,
                y = lat.W, label = ID),
            color = 'red',
            size = 3)+
  coord_map(xlim = c(-119, -117),
            ylim = c(32.5, 34)) +
  xlab('') + ylab('')

p.1

if (save.image) {
  ggsave(plot = p.1,
         dpi = 600,
         file = paste0('figures/tracklines_', 
                       km2nm(dlines/1000), 'nm_', Sys.Date(), '.png'))

  write.table(signif(survey.coords, digits = 8), sep = ',',
            file = paste0('data/tracklines_', 
                          km2nm(dlines/1000), 'nm_', Sys.Date(), '.csv'),
            append = F, quote = F, row.names = F, col.names = T)

  WP.file <- paste0("data/waypoints_",
                    km2nm(dlines/1000), 
                    'nm_', Sys.Date(), '.csv')
              
  write_csv(survey.coords.2, 
            file = WP.file)
  
  tmp.names <- read_csv(WP.file, 
                        col_types = cols(Pt = col_character(),
                                         Lat = col_double(),
                                         Lon = col_double())) %>%
    transmute(Pt = Pt,
              Lat = as.character(Lat),
              Lon = as.character(Lon)) %>% 
    apply(MARGIN = 1, 
          FUN = function(x){ paste0(x[1], ", ", 
                                    substr(x[2], 1, 7), ", ",
                                    substr(x[3], 1, 9))})
  
  survey.coords.2.sf <- st_as_sf(survey.coords.2, 
                                 coords = c("Lon", "Lat"))
  
  survey.coords.2.sp <- as(survey.coords.2.sf, "Spatial")
  maptools::kmlPoints(survey.coords.2.sp,
                      kmlfile = paste0("data/waypoints_",
                                       km2nm(dlines/1000), 
                                       'nm_', Sys.Date(), '.kml'),
                      description = "Track lines waypoints",
                      name = tmp.names,
                      icon = "http://maps.google.com/mapfiles/kml/shapes/placemark_circle.png")
  
  
}

```

