---
title: "Reading gpx files"
output: html_notebook
---

Trying to convert a gpx file into a dataframe. The survey on 22 August 2018 was conducted without having a laptop - no AC outlet on the plane... so, I recorded the track on my phone using iSailGPS app. The track was exported into a GPX file and emailed to myself. Now trying to convert it to sightings data as they would come out from carretta.

```{r}
library(rgdal)
library(dplyr)
library(lubridate)
ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))

```

Now try to read in the gpx file, change the zulu time into local time, and filter just one day's effort. This file contains all the old tracks from east coast also. 

```{r}
layers <- ogrListLayers("data/gpx files/2018-08-25_16-16-23 iSailGPS_Track.gpx")

# third layer is tracks
tracks_all <- readOGR(dsn = "data/gpx files/2018-08-25_16-16-23 iSailGPS_Track.gpx",
                      layer = "track_points")

# get time and elevation
data_time_elev <- tracks_all@data %>%
  select(ele, time)

# get the coordinates
data_coord <- tracks_all@coords

# combine them together and rename column names
data_coord_time_elev <- cbind(data_time_elev, data_coord)
colnames(data_coord_time_elev) <- c("Elev", "Date_Time", "Longitude", "Latitude")

# strptime returns POSIXlt, which is not filter-able below. So, use
# as.POSIXct.
data_coord_time_elev$Date_Z <- as.POSIXct(data_coord_time_elev$Date_Time,
                                          format = "%Y/%m/%d %H:%M:%S", 
                                          tz = "UTC") 
data_coord_time_elev$Date_local <- with_tz(data_coord_time_elev$Date_Z, 
                                           tzone = "America/Los_Angeles")

# filter out the most recent - today's
data_20180822 <- filter(data_coord_time_elev,
                        Date_local > as.POSIXct("2018-08-21 00:00:00", 
                                             format = "%Y-%m-%d %H:%M:%S")) %>%
  select(Date_local, Latitude, Longitude)

```

Need to read in sightings and on/off effort data here. Then combine with track line data.
```{r}
sightings.data <- read.csv(file = "data/Data_2018-08-22.csv",
                           header = TRUE) %>% # Convert date/time
  mutate(Date_local = as.POSIXct(paste(date, 
                                       time, 
                                       sep = " "),
                                 format = "%m/%d/%Y %H:%M:%S",
                                 tz = "America/Los_Angeles")) %>%
  select(Date_local, event, comment)

```

Need to add latitude/longitude to sightings data then sort by date/time.

```{r}
k <- 1
sightings.data$Latitude <- NA
sightings.data$Longitude <- NA
for (k in 1:nrow(sightings.data)){
  date.time <- sightings.data[k, "Date_local"]
  loc <- filter(data_20180822, Date_local >= date.time - 1 & Date_local <= date.time + 1)
  if (nrow(loc) == 1){
    sightings.data[k, c("Latitude", "Longitude")] <- loc[1, c("Latitude", "Longitude")]
  } else if (nrow(loc) > 1){
    sightings.data[k, c("Latitude", "Longitude")] <- colMeans(loc[, c("Latitude", "Longitude")])
  } 
}

data_20180822$event <- "*"
data_20180822$comment <- ""

sightings.data %>% select(event, Date_local, Latitude, Longitude, comment) -> sightings.data
data_20180822 %>% select(event, Date_local, Latitude, Longitude, comment) -> data_20180822

data_all <- rbind(sightings.data, data_20180822)
data_sorted <- data_all[order(data_all$Date_local),]
```

Then, write the data into a file. Break down each piece of information into the required format.

```{r}

sink("data/das files/SCB2018_DAS.822")
k<-1
c <- 1
on.effort <- F
time1 <- data_sorted[1, "Date_local"]
for (k in 1:nrow(data_sorted)){
  data.1 <- data_sorted[k,]
  event.code <- data.1$event
  
  if (event.code != "*"){
    line.number <- sprintf("%3d", c)
    c <- c + 1
    if (event.code == "T") on.effort <- T
    if (event.code == "O") on.effort <- F
    
    effort <- ifelse(on.effort, ".", " ")
    
    time <- paste0(formatC(hour(data.1$Date_local), width = 2, flag = 0),
                   formatC(minute(data.1$Date_local), width = 2, flag = 0),
                   formatC(second(data.1$Date_local), width = 2, flag = 0), " ")
    date <- paste0(formatC(month(data.1$Date_local), width = 2, flag = 0),
                   formatC(day(data.1$Date_local), width = 2, flag = 0),
                   "18 ")
    
    latitude <- paste0("N", floor(data.1$Latitude), ":",
                       formatC(((data.1$Latitude - floor(data.1$Latitude)) * 60), 
                               width = 5, flag = 0, format = "f", digits = 2), " ")
    
    longitude <- paste0("W", floor(abs(data.1$Longitude)), ":",
                        formatC(((abs(data.1$Longitude) - floor(abs(data.1$Longitude))) * 60), 
                                width = 5, flag = 0, format = "f", digits = 2), " ")
    
    if (event.code == "T"){
      str.data <- data.1$comment
    } else if (event.code == "V"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), 
                                          ";")), width = 5)
    } else if (event.code == "P"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "A"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "W") {
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "S"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "C"){
      str.data <- data.1$comment
    } else {
      str.data <- ""
    }
    out.line <- paste0(line.number, event.code, effort,
                     time, date, latitude, longitude, 
                     paste(str.data, sep = "", collapse = ""),
                     "\n")
    cat(out.line)
  } else {
    time2 <- data_sorted[k, "Date_local"]
    dt <- time2 - time1
    if (dt > duration(minute = 1)){
      line.number <- sprintf("%3d", c)
      time1 <- time2
      effort <- ifelse(on.effort, ".", " ")
      
      time <- paste0(formatC(hour(data.1$Date_local), width = 2, flag = 0),
                     formatC(minute(data.1$Date_local), width = 2, flag = 0),
                     formatC(second(data.1$Date_local), width = 2, flag = 0), " ")
      date <- paste0(formatC(month(data.1$Date_local), width = 2, flag = 0),
                     formatC(day(data.1$Date_local), width = 2, flag = 0),
                     "18 ")
      
      latitude <- paste0("N", floor(data.1$Latitude), ":",
                         formatC(((data.1$Latitude - floor(data.1$Latitude)) * 60), 
                                 width = 5, flag = 0, format = "f", digits = 2), " ")
      
      longitude <- paste0("W", floor(abs(data.1$Longitude)), ":",
                          formatC(((abs(data.1$Longitude) - floor(abs(data.1$Longitude))) * 60), 
                                  width = 5, flag = 0, format = "f", digits = 2), " ")
      str.data <- ""
      c <- c + 1
      out.line <- paste0(line.number, event.code, effort,
                     time, date, latitude, longitude, 
                     paste(str.data, sep = "", collapse = ""),
                     "\n")
      cat(out.line)
    }
  }

}
sink()



```

