---
title: "Reading gpx files"
output: html_notebook
---

Fixing DAS file for 28 August 2018 because the GPS was not functioning properly between 12:47:00 and 12:53:19.

```{r}
rm(list=ls())
library(rgdal)
library(dplyr)
library(lubridate)
ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))

```

Now try to read in the gpx file, change the zulu time into local time, and filter just one day's effort. This file contains all the old tracks from east coast also. 

```{r}
layers <- ogrListLayers("Data/gpx files/2018-08-28_15-22-02 iSailGPS_Track.gpx")

# third layer is tracks
tracks_all <- readOGR(dsn = "Data/gpx files/2018-08-28_15-22-02 iSailGPS_Track.gpx",
                      layer = "track_points")

# get time and elevation
data_time_elev <- tracks_all@data %>%
  select(ele, time)

# get the coordinates
data_coord <- tracks_all@coords

# combine them together and rename column names
data_coord_time_elev <- cbind(data_time_elev, data_coord)
colnames(data_coord_time_elev) <- c("Elev", "Date_Time", "Longitude", "Latitude")

# strptime returns POSIXlt, which is not filter-able below. So, use
# as.POSIXct.
data_coord_time_elev$Date_Z <- as.POSIXct(data_coord_time_elev$Date_Time,
                                          format = "%Y/%m/%d %H:%M:%S", 
                                          tz = "UTC") 
data_coord_time_elev$Date_local <- with_tz(data_coord_time_elev$Date_Z, 
                                           tzone = "America/Los_Angeles")

# took off around 12:27
data_20180828 <- filter(data_coord_time_elev,
                        Date_local > as.POSIXct("2018-08-28 12:46:38", 
                                             format = "%Y-%m-%d %H:%M:%S") &
                        Date_local < as.POSIXct("2018-08-28 12:53:19", 
                                             format = "%Y-%m-%d %H:%M:%S")  ) %>%
  select(Date_local, Latitude, Longitude)

```

Check what we missed - date-time combos are...
```{r}
missed.time <- c("2018-08-28 12:48:00",
                 "2018-08-28 12:48:40",
                 "2018-08-28 12:48:53",
                 "2018-08-28 12:49:00",
                 "2018-08-28 12:49:44",
                 "2018-08-28 12:50:00",
                 "2018-08-28 12:50:43",
                 "2018-08-28 12:50:58",
                 "2018-08-28 12:51:00",
                 "2018-08-28 12:52:00",
                 "2018-08-28 12:53:00",
                 "2018-08-28 12:53:07")

lat.deg <- lat.min <- vector(mode = "numeric", length = length(missed.time))
lon.deg <- lon.min <- vector(mode = "numeric", length = length(missed.time))

k <- 1
for (k in 1:length(missed.time)){
  time1 <- as.POSIXct(missed.time[k],format = "%Y-%m-%d %H:%M:%S") - duration(2, units = "seconds")
  time2 <- as.POSIXct(missed.time[k],format = "%Y-%m-%d %H:%M:%S") + duration(2, units = "seconds")
  data_20180828 %>% filter(Date_local > time1 & Date_local < time2) -> tmp
  
    lat.deg[k] <- floor(tmp$Latitude[1])
    lat.min[k] <- (tmp$Latitude[1] - lat.deg[k])*60
    lon.deg[k] <- floor(abs(tmp$Longitude[1]))
    lon.min[k] <- (abs(tmp$Longitude[1]) - lon.deg[k])*60
    
}


```

Need to add latitude/longitude to sightings data then sort by date/time.

```{r}
# also another a few lines...
data_20180828 <- filter(data_coord_time_elev,
                        Date_local > as.POSIXct("2018-08-28 14:19:18", 
                                             format = "%Y-%m-%d %H:%M:%S") &
                        Date_local < as.POSIXct("2018-08-28 14:20:40", 
                                             format = "%Y-%m-%d %H:%M:%S")  ) %>%
  select(Date_local, Latitude, Longitude)

missed.time <- c("2018-08-28 14:20:00",
                 "2018-08-28 14:20:24")

lat.deg <- lat.min <- vector(mode = "numeric", length = length(missed.time))
lon.deg <- lon.min <- vector(mode = "numeric", length = length(missed.time))

k <- 1
for (k in 1:length(missed.time)){
  time1 <- as.POSIXct(missed.time[k],format = "%Y-%m-%d %H:%M:%S") - duration(2, units = "seconds")
  time2 <- as.POSIXct(missed.time[k],format = "%Y-%m-%d %H:%M:%S") + duration(2, units = "seconds")
  data_20180828 %>% filter(Date_local > time1 & Date_local < time2) -> tmp
  
  lat.deg[k] <- floor(tmp$Latitude[1])
  lat.min[k] <- (tmp$Latitude[1] - lat.deg[k])*60
  lon.deg[k] <- floor(abs(tmp$Longitude[1]))
  lon.min[k] <- (abs(tmp$Longitude[1]) - lon.deg[k])*60
    
}
```

Then, write the data into a file. Break down each piece of information into the required format.

```{r}

sink("data/das files/SCB2018_DAS.826")
k<-1
c <- 1
on.effort <- F
time1 <- data_sorted[1, "Date_local"]
for (k in 1:nrow(data_sorted)){
  data.1 <- data_sorted[k,]
  event.code <- data.1$event
  
  if (event.code != "*"){
    line.number <- sprintf("%3d", c)
    c <- c + 1
    if (event.code == "T" | event.code == "R") on.effort <- T
    if (event.code == "O" | event.code == "E") on.effort <- F
    
    effort <- ifelse(on.effort, ".", " ")
    
    time <- paste0(formatC(hour(data.1$Date_local), width = 2, flag = 0),
                   formatC(minute(data.1$Date_local), width = 2, flag = 0),
                   formatC(second(data.1$Date_local), width = 2, flag = 0), " ")
    date <- paste0(formatC(month(data.1$Date_local), width = 2, flag = 0),
                   formatC(day(data.1$Date_local), width = 2, flag = 0),
                   "18 ")
    
    latitude <- paste0("N", floor(data.1$Latitude), ":",
                       formatC(((data.1$Latitude - floor(data.1$Latitude)) * 60), 
                               width = 5, flag = 0, format = "f", digits = 2), " ")
    
    longitude <- paste0("W", floor(abs(data.1$Longitude)), ":",
                        formatC(((abs(data.1$Longitude) - floor(abs(data.1$Longitude))) * 60), 
                                width = 5, flag = 0, format = "f", digits = 2))
    
    if (event.code == "T"){
      str.data <- formatC(data.1$comment, width = 5)
    } else if (event.code == "V"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), 
                                          ";")), width = 5)
    } else if (event.code == "P"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "A"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "W") {
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "S"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else if (event.code == "C"){
      str.data <- paste0(" ", data.1$comment)
    } else if (event.code == "t"){
      str.data <- formatC(unlist(strsplit(as.character(data.1$comment), ";")), 
                          width = 5)
    } else {
      str.data <- ""
    }
    out.line <- paste0(line.number, event.code, effort,
                       time, date, latitude, longitude, 
                       paste(str.data, sep = "", collapse = ""),
                       "\n")
    cat(out.line)
  } else {
    time2 <- data_sorted[k, "Date_local"]
    dt <- time2 - time1
    if (dt > duration(minute = 1)){
      line.number <- sprintf("%3d", c)
      time1 <- time2
      effort <- ifelse(on.effort, ".", " ")
      
      time <- paste0(formatC(hour(data.1$Date_local), width = 2, flag = 0),
                     formatC(minute(data.1$Date_local), width = 2, flag = 0),
                     formatC(second(data.1$Date_local), width = 2, flag = 0), " ")
      date <- paste0(formatC(month(data.1$Date_local), width = 2, flag = 0),
                     formatC(day(data.1$Date_local), width = 2, flag = 0),
                     "18 ")
      
      latitude <- paste0("N", floor(data.1$Latitude), ":",
                         formatC(((data.1$Latitude - floor(data.1$Latitude)) * 60), 
                                 width = 5, flag = 0, format = "f", digits = 2), " ")
      
      longitude <- paste0("W", floor(abs(data.1$Longitude)), ":",
                          formatC(((abs(data.1$Longitude) - floor(abs(data.1$Longitude))) * 60), 
                                  width = 5, flag = 0, format = "f", digits = 2))
      str.data <- ""
      c <- c + 1
      out.line <- paste0(line.number, event.code, effort,
                     time, date, latitude, longitude, 
                     paste(str.data, sep = "", collapse = ""),
                     "\n")
      cat(out.line)
    }
  }

}
sink()

```


